---
title: First Analysis
output: 
  # tufte::tufte_html
  html_document:
    code_folding: hide
    css: styles.css
---
```{r, include=FALSE}
# Used to load the data
source("load_data.R")
source("functions.R")
knitr::opts_chunk$set(fig.width=14, fig.height=7, dpi=300, out.width="100%")
knitr::opts_knit$set(root.dir = normalizePath("."))
```


## The data set
Data for all the Isar stations from GKD website. 

```{r}
cop_df = get_copula_df()
```

*Exclude* station Moosburg because only have $5$ data points here. Aparently the data collection process still young (first year is 2020)?
```{r}
# Moosburg only has 5 observations anyway. Think I should drop it
cop_df |> dplyr::filter(unit == "Moosburg") 
# Dropping Moosburg:
cop_df = cop_df |> dplyr::filter(unit != "Moosburg")
```
Data after dropping Moosburg:
```{r}
cop_df = cop_df |> 
  dplyr::mutate(
    pobs_dur = copula::pobs(duration_min),
    pobs_peak = copula::pobs(peak),
    pobs_vol= copula::pobs(volume),
    .by = unit
  )
knitr::kable(head(cop_df))
```


## Aggregated view
Scatterplots of observed variables (Marginal affects Scatterplot) and Pseudo-Observations (Relationship only).

### Volume - Duration
```{r}
# Plots
# Volume-Duration
# With marginal distribution 
p1 = ggplot(cop_df, aes(x = volume, y = duration_min)) + 
  geom_point() + 
  facet_wrap(~unit)
# Pobs
p2 = ggplot(cop_df, aes(x = pobs_vol, y = pobs_dur)) + 
  geom_point() + 
  facet_wrap(~unit)
p1p2 = (p1 | p2)
p1p2
```

### Volume - Peak
```{r}
# Volume-Peak
# W/ marginal
p1 = ggplot(cop_df, aes(x = volume, y = peak)) + 
  geom_point() + 
  facet_wrap(~unit)
# Pobs (relation without marginal)
p2 = ggplot(cop_df, aes(x = pobs_vol, y = pobs_peak)) + 
  geom_point() + 
  facet_wrap(~unit)
p1p2 = (p1 | p2)
plot(p1p2)
```


### Duration - Peak
```{r}
# Duration-Peak
# W/ marginal
p1 = ggplot(cop_df, aes(x = duration_min, y = peak)) + 
  geom_point() + 
  facet_wrap(~unit)
# Pobs
p2 = ggplot(cop_df, aes(x = pobs_dur, y = pobs_peak)) + 
  geom_point() + 
  facet_wrap(~unit)
p1p2 = (p1|p2)
plot(p1p2)
```


## Correlation table
```{r}
# Correlation table
cor_table = cop_df |> dplyr::summarise(
  n = dplyr::n(),
  tau_vd = cor(volume, duration_min, method = "kendall"),
  p_vd = cor.test(volume, duration_min, method = "kendall")$p.value,
  rej_vd = p_vd < 0.01,
  tau_vp = cor(volume, peak, method = "kendall"),
  p_vp = cor.test(volume, peak, method = "kendall")$p.value,
  rej_vp = p_vp < 0.01,
  tau_dp = cor(duration_min, peak, method = "kendall"),
  p_dp = cor.test(duration_min, peak, method = "kendall")$p.value,
  rej_dp = p_dp < 0.01,
  .by = unit
)
knitr::kable(cor_table)
```


#### Coorelation plots

For NACs, we only model $d-1$ relations. That is, the two strongly dependent variables have the same (high) $\tau$, and the weaker variable should have the same (low) degree of dependence wrt. the other two variables. This does not look like it tho:\
For NACs to be applicable, I expected two very similar Boxplots and one with larger values. This looks like I would have to model 3 separate relations...\

```{r}
# Distribution of dependence structures
cor_table |> tidyr::pivot_longer(
  cols = c("tau_vd", "tau_vp", "tau_dp"),
  names_to = "relation",
  values_to = "correlation"
) |>
ggplot(aes(y = correlation, x = relation)) + 
  geom_boxplot()
```
(tau_vd): I mean, volume and duration are by nature highly correlated due to their functional dependence. After all, if duration increases and the discharge is greater than 0, the volume increases too.\
(tau_vp): For volume and peak, there exists some relation? It seems like with increasing peak, the flood volume increases which is reasonable too since for larger peak, I expect generally larger discharge values (could check this in the data). So generally larger discharge translates to larger volume.\
(tau_dp) Duration and peak only has a weak correlation strucutre. Implying that even for large peaks, the duration need not to be long.

So basically, I have $3$ different relations, but NACs only allow us to model $2$ separate relations. lol. Is this why the paper never showed a scatter plot of their pobs? 

So back to Vines vs. NACs? Could run a nice simulation. Also, I can list the limitations of NACs, not only the relation part. Also the estimation process.

OKAAY: Since NACs are overall just super annoying in modelling and, well, overall, idea is:\
1) Simulation study to show how bad NACs are if their conditions are not fulfilled. Here I can also mention that NAC-R-packages mainly focus on the nested structure of one single family. Thus, R packages for this family are super limited anyway which implies a bunch of self implementation if NACs want to be used.\
Showing NACs are bad: Simulate with different n, dependence tau (tail dependence lambda), ?, and fit vine vs NAC and compare fits using SOMETHING? At best a measure that is independent of the likelihood to ensure that I am able to not gaf about it.\
2) Fit both models to the data and simulate based on fitted. Show that simulated data is shit for NACs bc assumptions not valid. (Or maybe show that NACs go brr even tho assumptions totally do not hold. lol. Doubt it.)










